---
title: <span style="color:white">The Displacement Effect</span>
subtitle: "Exploring the effect in Sub-Sarahan African (SSA) Countries (2000-2022)"

author: 
- name: "Matthew Kuch"
  email: matt@bk-advisors.com
date: 08/12/2025

title-block-banner: true
title-block-banner-color: "black"

format: html
html:
code-fold: true
code-summary: "Code"
echo: true
css: style.css
toc: true
toc-location: left
number-sections: false
editor: visual
fig-cap-location: margin
---

## The Spark of Curiosity

There's probably something many don't know about me; I have a peculiar past-time or hobby, if you like.

Whenever I come across a new idea or hypothesis or theory - I try to look for available data to test it, just out of curiosity, to see the limits of the theory or edge cases.

Same it true for the "displacement effect", which I wrote about this week on Linkedin.

In this more technical post, I'll walk you through my analysis of the WHO GHED (Global Health Expenditure Data) from 2000 to 2022 to test the claim I made in the article, that:

>  *"an increase is funding to public health from external donors, leads to the government spending less on public health."*

## The Core Question

The core question I'll be asking is this:

> “If donors give more money to a country’s health budget one year, does the government spend less of its own money the next year?”

Like I mentioned in my article, its kind of like if your child received a partial scholarship as a contribution to their school fees, for say good performance, but instead of stepping up as a parent and paying the balance - you instead stop paying school fees altogether and use the money you had to buy the new iPhone.\

That’s the **displacement effect** - when outside help replaces rather than adds to your own effort.

```{r}
1 + 1
```

You can add options to executable code like this

```{r}
#| echo: false
2 * 2
```

The `echo: false` option disables the printing of code (only output is displayed).
